---
title: "scMOO-tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{scMOO-tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## 1. Installation
Step 1. If the "devtools" package has not been installed, install the "devtools" package first.
```{r setup, eval=FALSE}
install.packages("devtools")
```
Step 2. Install the "scMOO" package from GitHub.
```{r install, eval=FALSE}
devtools::install_github("Zhangxf-ccnu/scMOO", subdir="pkg")
```
## 2. Quick Start
To perform imputation, simply run:
```{r scmoo, eval=FALSE}
library(scMOO)
data(PBMC_CL)
result <- scMOO(PBMC_CL, percent=0, verbose=TRUE, estimates.only=FALSE)
```
Here we use PBMC_CL dataset used in the cell clustering analysis as an example. Note that the PBMC_CL dataset has been preporcessed using Seurat with 2,000 selected highly variable genes and whose column names are true cell labels.  
&nbsp;&nbsp;&nbsp;&nbsp;`percent`: The expression count matrix is preprocessed by filtering out the genes expressed in at most percent*100\% of the cells. In the above example, we set `percent=0` for that we have selected 2,000 highly variable genes.    
&nbsp;&nbsp;&nbsp;&nbsp;`verbose`: Whether to output the value of metrics at the end of each epoch. Default is TRUE.   
&nbsp;&nbsp;&nbsp;&nbsp;`estimates.only`: If `estimates.only = TRUE`, `scMOO()` returns a matrix of scMOO estimates;   
 &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;  &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; &nbsp;&nbsp;&nbsp;&nbsp;                  If `estimates.only = FALSE`, `scMOO()` returns the following list of values:      
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;  &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; `estimate`: Recovered (normalized) expression.   
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;  &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; `size.factor`: Size factor used for normalization.   
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;  &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; `pred.time`: Total time for scMOO estimate.   
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;  &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; `alpha`: Tuning parameter to balance the error between the imputed and observed data &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;  &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;and the eror of model fitting data.    
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;  &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; `w`: The combination weights $\omega_{1}$, $\omega_{2}$ and $\omega_{3}$.

## 3. Downstream analysis: Cell clustering 
### 3.1 Selecting highly variable genes before imputation
```{r hvgs, eval=FALSE}
library(Seurat)
x.seurat <- CreateSeuratObject(raw_data) ## colnames of data are ture cell labels
x.seurat <- NormalizeData(x.seurat)
x.seurat <- ScaleData(x.seurat)
x.seurat <- FindVariableFeatures(x.seurat, verbose = FALSE)

filter_ID <- x.seurat@assays$RNA@var.features
data_hvgs <- data[filter_ID,]
```
Through 3.1, we can obtain PBMC_CL dataset with 2,000 highly variable genes before imputation.      

### 3.2 Imputing dropouts using scMOO
```{r imputation, eval=FALSE}
library(scMOO)
result <- scMOO(PBMC_CL,  percent=0, estimate.only=TRUE)
```
By setting `estimate.only=TRUE`, we can obtain a matrix of scMOO estimates.

### 3.3 Performing cell clustering using SC3
```{r sc3, eval=FALSE}
library(SingleCellExperiment)
library(SC3)
library(mclust)
library(aricode)

sce <- SingleCellExperiment(
  assays = list(
    counts = as.matrix(result),
    logcounts = log2(as.matrix(result) + 1)
  ), 
  colData = colnames(result)
)


rowData(sce)$feature_symbol <- rownames(sce)

label <- colnames(result)

K <- length(unique(label))


result <- sc3(sce, ks = K, gene_filter = FALSE) 

ARI <- adjustedRandIndex(label, result@colData@listData$sc3_6_clusters) #### '6' in 'sc3_6_clusters' is consistant, and equals to the number of cell types 

NMI <- NMI(result@colData@listData$sc3_6_clusters, label)

```
We first perform cell lustering using `SC3` on the imputed results. The R package
`SC3` is for implementing `SC3`, and the parameter `ks`, which determines
the number of clusters, is chosen as the known number of cell types. Here `ks` is set as 6 on the PBMC_CL dataset.  
To compare the clustering results, we use the **Adjusted Rand Index**
(ARI) and the **Normalized Mutual Information** (NMI) to evaluate the
consistency between the results of `SC3` and the reference labels
of cells. The closer the indexes of **ARI** and **NMI** are to 1, the better the
clustering results will be.

### 3.4 Performing cell clustering using Seurat
```{r seurat, eval=FALSE}
library(Seurat)
x.seurat <- CreateSeuratObject(result) 
x.seurat <- NormalizeData(x.seurat)
x.seurat <- ScaleData(x.seurat)
x.seurat <- FindVariableFeatures(x.seurat, verbose = FALSE)

x.seurat <- RunPCA(x.seurat, features = VariableFeatures(object = x.seurat))  
x.seurat <- JackStraw(x.seurat, num.replicate = 100)
x.seurat <- ScoreJackStraw(x.seurat, dims = 1:20)
x.seurat <- FindNeighbors(x.seurat, dims = 1:10)
x.seurat <- FindClusters(x.seurat, resolution = 0.5)

label <- colnames(result)

ARI <- adjustedRandIndex(as.factor(label), x.seurat$seurat_clusters)

NMI <- NMI(x.seurat$seurat_clusters, as.factor(label))
```
We then perform cell lustering using `Seurat` on the imputed results. Similarly, **ARI** and **NMI** are used to evaluate the consistency between the results of `Seurat` and the reference labels of cells.
